---
title: "RF_Nekton_Habitat_Suitability"
output: html_document
date: "2026-04-01"
---
## Author: Jon Rodemann

```{r setup, include=FALSE}
library(tidyverse)
library(randomForest)
library(mlr3verse)
library(iml)
library(ranger)
```

The following code goes through various iterations of random forest models to look at habitat suitability of nekton in seagrass habitats

```{r, include=FALSE}
#Load in datasets
sled <- read.csv('Data/combined_sled_species_sp.csv')
trawl <- read.csv('Data/combined_trawl_species.csv')

##We want to isolate the 8 important fish species. Here are the ones:
# - Gray snapper
# - Blue crab
# - Pink shrimp
# - Spotted seatrout
# - Red drum
# - White shrimp
# - Brown shrimp
# - Lane snapper

##Ok. We are going to try a couple of things. Here are the different models we are going to try:
#Dependent variable:
# Binary
# Abundance with zeros
# Hurdle RF
#Independent variables:
# All
# Handpicked set
```

## Binary with all variables

This model will translate the abundance data into presence/absence data and then train a RF model for binary on all of the independent variables for each species. Will do sled and trawl separately and then together.

First let's organize the raw data into something we can use

```{r}
#First we have to organize the data. This will include isolating only the species of interest and including all of the zeros
#add in all zeros
#Extract all of the station, site, and sample_period data to combine it together after
stat_dat_sled <- sled %>% dplyr::select(site, station, sample_period, light_attenuation_av:pull_id, station_seagrass_richness:sf_station_sd_below_wt_m) %>% distinct()

stat_dat_trawl <- trawl %>% dplyr::select(tow_id:station_sd_rm_sht_ratio, site, station, sample_period, tow_area:light_attenuation_av, seagrass_richness) %>% distinct()

#add in all zeros
sled_zero <- sled %>% dplyr::select(site, station, sample_period, species_common, species_abundance) %>% 
  dplyr::group_by(site, station, sample_period, species_common) %>% 
  dplyr::summarise(species_abundance = sum(species_abundance, na.rm = TRUE), .groups = "drop") %>%
  # Use complete() to generate all combinations
  complete(
    nesting(site, station, sample_period), # Only crosses combinations that actually exist in your data
    species_common,                        # Crosses the above with all unique species
    fill = list(species_abundance = 0)     # Fills implicitly missing abundances with 0
  )

trawl_zero <- trawl %>% dplyr::select(site, station, sample_period, species_common, species_abundance) %>% 
  dplyr::group_by(site, station, sample_period, species_common) %>% 
  dplyr::summarise(species_abundance = sum(species_abundance, na.rm = TRUE), .groups = "drop") %>%
  # Use complete() to generate all combinations
  complete(
    nesting(site, station, sample_period), # Only crosses combinations that actually exist in your data
    species_common,                        # Crosses the above with all unique species
    fill = list(species_abundance = 0)     # Fills implicitly missing abundances with 0
  )

#recombine the data
sledall <- sled_zero %>% left_join(stat_dat_sled, by = c('site', 'station', 'sample_period'))
trawlall <- trawl_zero %>% left_join(stat_dat_trawl, by = c('site', 'station', 'sample_period'))

#Isolate species of interest and replace NAs with zeros
sledsp <- sledall %>% dplyr::filter(species_common %in% c('gray snapper', 'blue crab', 'pink shrimp', 'white shrimp', 'brown shrimp', 'spotted seatrout', 'lane snapper', 'red drum')) %>% mutate(across(where(is.numeric), ~replace_na(.x, 0))) %>% dplyr::select(-c(spp_x, hd_station_ep_wt_gg, hd_station_sd_ep_wt_gg, hd_station_below_wt_m, hd_station_sd_below_wt_m, spp_y, he_station_sd_ep_wt_gg, he_station_sd_below_wt_m, rm_stationsp_sd_shoots, rm_stationsp_sd_shoots_m, rm_stationsp_se_shoots_m, rm_station_sd_above_wt_m, rm_station_ep_wt_gg, rm_station_sd_ep_wt_gg, rm_station_below_wt_m, rm_station_sd_below_wt_m))

trawlsp <- trawlall %>% dplyr::filter(species_common %in% c('gray snapper', 'blue crab', 'pink shrimp', 'white shrimp', 'brown shrimp', 'spotted seatrout', 'lane snapper', 'red drum')) %>% mutate(across(where(is.numeric), ~replace_na(.x, 0)))

#Ok. We have our dataset with abundances and zeros. We are going to add a column to indicate binary presence/absence and an ID column
sledbin <- sledsp %>% mutate(pa = if_else(species_abundance == 0, 0, 1), ID = row_number())

trawlbin <- trawlsp %>% mutate(pa = if_else(species_abundance == 0, 0, 1), ID = row_number())

#combine the trawl and sled datasets for a more complete dataset
sledbin <- sledbin %>% mutate(type = 'sled')
trawlbin <- trawlbin %>% mutate(type = 'trawl')

st <- bind_rows(sledbin, trawlbin) %>% group_by(site, station, sample_period) %>% fill(everything(), .direction = 'downup') %>% ungroup()
```

We set up the binary data for both datasets. Let's set up the RF binary model

```{r Binary_brown_shrimp}
#filter for brown shrimp
data <- st %>% filter(species_common == 'brown shrimp')

data$pa <- as.factor(data$pa)

#Take out columns that will not be in model
data <- data %>% dplyr::select(-c(station, sample_period, species_common, species_abundance, secchi, pull_id, site))

str(data)
data$depth <- as.numeric(data$depth)
data$station_seagrass_richness <- as.numeric(data$station_seagrass_richness)
set.seed(19)
# Randomly select 70% of the data frame for the training dataset - from each
datasled <- data %>% dplyr::filter(type == "sled")
datatr <- data %>% dplyr::filter(type == 'trawl')
data.trains <- datasled[sample(1:nrow(datasled), nrow(datasled) * 0.8, replace = FALSE), ]
data.traint <- datatr[sample(1:nrow(datatr), nrow(datatr) * 0.8, replace = FALSE), ]
# Remainder (i.e., 30%) becomes the test dataset.
data.tests <- datasled[!(datasled$ID %in% data.trains$ID), ]
data.testt <- datatr[!(datatr$ID %in% data.traint$ID), ]


data.train <- bind_rows(data.trains, data.traint) %>% dplyr::select(-c(ID, tow_id))
data.test <- bind_rows(data.tests, data.testt) %>% dplyr::select(-c(ID, tow_id))

#remove columns with Inf data
data.train <- data.train %>% dplyr::select(where(~ !is.numeric(.x) || !any(is.infinite(.x))))
data.test <- data.test %>% dplyr::select(where(~ !is.numeric(.x) || !any(is.infinite(.x))))

#change columns with NA data to zeros
data.train <- data.train %>% dplyr::mutate(across(where(is.numeric), ~replace_na(.x, 0)))
data.test <- data.test %>% dplyr::mutate(across(where(is.numeric), ~replace_na(.x, 0)))

#set tasks for training and test datasets
task_train <- as_task_classif(x = data.train, target = 'pa', positive = '1')
task_test <- as_task_classif(x = data.test, target = 'pa', positive = '1')

# Make learner.
learner <-lrn(
  "classif.ranger",
  predict_type = "prob",
  mtry  = to_tune(1, ncol(data.train) - 4),
  sample.fraction = to_tune(0.2, 0.9),
  min.node.size = to_tune(p_int(1,10)),
  importance = 'impurity'
)

#tune hyperparameters
instance = ti(
  task = task_train,
  learner = learner,
  resampling = rsmp("cv", folds = 5),
  measures = msr("classif.ce"),
  terminator = trm("none")
)

tuner = mlr3tuning::tnr("grid_search", resolution = 5, batch_size = 5)
tuner$optimize(instance) # Takes ~ 4 minutes on my relatively fast computer

#store tuned hyperparameters in learner
learner$param_set$values <- instance$result_learner_param_vals

#finally! We can train our model with the train function
learner$train(task_train)

#let's quickly look at the model
learner$model

#Accuracy of model - first on training data, then testing data
measures <- msrs(c('classif.acc'))

pred_train <- learner$predict(task_train)
pred_train$confusion
pred_train$score(measures)


pred_test <- learner$predict(task_test)
pred_test$confusion
pred_test$score(measures)

#importance with iml package - this is looking at the most influencial predictors in the model
x_data <- data.train %>% dplyr::select(-pa) 
# Create "Predictor" object to interpret findings via the iml package.
predictor_data <- Predictor$new(learner, data = x_data, y = data.train$pa) 

imp_data <- FeatureImp$new(predictor_data, loss = "ce") # Calculate importance.

imp_df_data <- imp_data$results 


#most important covariate is the type of collection. Here is what it looks like
effect_type <- FeatureEffect$new(predictor_data, feature = c('type'), method = 'pdp')
typebrownshrimp <- effect_type$plot()

#so salinity is the most important variable for brown shrimp out of the environmental data. Let's investigate how it impacts the presence of shrimp

effect_sal <- FeatureEffect$new(predictor_data, feature = c('salinity'), method = 'pdp')
effect_sal$plot()
salbrownshrimp <- effect_sal$plot()

#temperature is the next important
effect_temp <- FeatureEffect$new(predictor_data, feature = c('temp'), method = 'pdp')
effect_temp$plot()
tempbrownshrimp <- effect_temp$plot()

#next are 2 standard errors of canopy height! One for overall, one for thalassia
effect_chse <- FeatureEffect$new(predictor_data, feature = c('station_se_mcanopy'), method = 'pdp')
effect_chse$plot()
chsebrownshrimp <- effect_chse$plot()


effect_chsett <- FeatureEffect$new(predictor_data, feature = c('tt_stationsp_se_mcanopy'), method = 'pdp')
effect_chsett$plot()
chsettbrownshrimp <- effect_chsett$plot()

#Overall, lower probability of presence at lower salinities, higher temperatures, and lower se of canopy height. More likely to catch brown shrimp in sled
```


```{r Binary_spotted_seatrout}
#filter for spotted seatrout
data <- st %>% filter(species_common == 'spotted seatrout')

data$pa <- as.factor(data$pa)
inv <- as.data.frame(data$pa)

#Take out columns that will not be in model
data <- data %>% dplyr::select(-c(station, sample_period, species_common, species_abundance, secchi, pull_id, site))

str(data)
data$depth <- as.numeric(data$depth)
data$station_seagrass_richness <- as.numeric(data$station_seagrass_richness)
set.seed(19)
# Randomly select 70% of the data frame for the training dataset - from each
datasled <- data %>% dplyr::filter(type == "sled")
datatr <- data %>% dplyr::filter(type == 'trawl')
data.trains <- datasled[sample(1:nrow(datasled), nrow(datasled) * 0.8, replace = FALSE), ]
data.traint <- datatr[sample(1:nrow(datatr), nrow(datatr) * 0.8, replace = FALSE), ]
# Remainder (i.e., 30%) becomes the test dataset.
data.tests <- datasled[!(datasled$ID %in% data.trains$ID), ]
data.testt <- datatr[!(datatr$ID %in% data.traint$ID), ]


data.train <- bind_rows(data.trains, data.traint) %>% dplyr::select(-c(ID, tow_id))
data.test <- bind_rows(data.tests, data.testt) %>% dplyr::select(-c(ID, tow_id))

#remove columns with Inf data
data.train <- data.train %>% dplyr::select(where(~ !is.numeric(.x) || !any(is.infinite(.x))))
data.test <- data.test %>% dplyr::select(where(~ !is.numeric(.x) || !any(is.infinite(.x))))

#change columns with NA data to zeros
data.train <- data.train %>% dplyr::mutate(across(where(is.numeric), ~replace_na(.x, 0)))
data.test <- data.test %>% dplyr::mutate(across(where(is.numeric), ~replace_na(.x, 0)))

#set tasks for training and test datasets
task_train <- as_task_classif(x = data.train, target = 'pa', positive = '1')
task_test <- as_task_classif(x = data.test, target = 'pa', positive = '1')

# Make learner.
learner <-lrn(
  "classif.ranger",
  predict_type = "prob",
  mtry  = to_tune(1, ncol(data.train) - 4),
  sample.fraction = to_tune(0.2, 0.9),
  min.node.size = to_tune(p_int(1,10)),
  importance = 'impurity'
)

#tune hyperparameters
instance = ti(
  task = task_train,
  learner = learner,
  resampling = rsmp("cv", folds = 5),
  measures = msr("classif.ce"),
  terminator = trm("none")
)

tuner = mlr3tuning::tnr("grid_search", resolution = 5, batch_size = 5)
tuner$optimize(instance) # Takes ~ 4 minutes on my relatively fast computer

#store tuned hyperparameters in learner
learner$param_set$values <- instance$result_learner_param_vals

#finally! We can train our model with the train function
learner$train(task_train)

#let's quickly look at the model
learner$model

#Accuracy of model - first on training data, then testing data
measures <- msrs(c('classif.acc'))

pred_train <- learner$predict(task_train)
pred_train$confusion
pred_train$score(measures)


pred_test <- learner$predict(task_test)
pred_test$confusion
pred_test$score(measures)

#importance with iml package - this is looking at the most influencial predictors in the model
x_data <- data.train %>% dplyr::select(-pa) 
# Create "Predictor" object to interpret findings via the iml package.
predictor_data <- Predictor$new(learner, data = x_data, y = data.train$pa) 

imp_data <- FeatureImp$new(predictor_data, loss = "ce") # Calculate importance.

imp_df_data <- imp_data$results 

#The random forest model does not work for the seatrout. It thinks that there are no seatrout. I don't know why, maybe will change when we pick and choose parameters
```

```{r Binary_lane_snapper}
#filter for lane snapper
data <- st %>% filter(species_common == 'lane snapper')

data$pa <- as.factor(data$pa)
inv <- as.data.frame(data$pa)

#Take out columns that will not be in model
data <- data %>% dplyr::select(-c(station, sample_period, species_common, species_abundance, secchi, pull_id, site))

str(data)
data$depth <- as.numeric(data$depth)
data$station_seagrass_richness <- as.numeric(data$station_seagrass_richness)
set.seed(19)
# Randomly select 70% of the data frame for the training dataset - from each
datasled <- data %>% dplyr::filter(type == "sled")
datatr <- data %>% dplyr::filter(type == 'trawl')
data.trains <- datasled[sample(1:nrow(datasled), nrow(datasled) * 0.8, replace = FALSE), ]
data.traint <- datatr[sample(1:nrow(datatr), nrow(datatr) * 0.8, replace = FALSE), ]
# Remainder (i.e., 30%) becomes the test dataset.
data.tests <- datasled[!(datasled$ID %in% data.trains$ID), ]
data.testt <- datatr[!(datatr$ID %in% data.traint$ID), ]


data.train <- bind_rows(data.trains, data.traint) %>% dplyr::select(-c(ID, tow_id))
data.test <- bind_rows(data.tests, data.testt) %>% dplyr::select(-c(ID, tow_id))

#remove columns with Inf data
data.train <- data.train %>% dplyr::select(where(~ !is.numeric(.x) || !any(is.infinite(.x))))
data.test <- data.test %>% dplyr::select(where(~ !is.numeric(.x) || !any(is.infinite(.x))))

#change columns with NA data to zeros
data.train <- data.train %>% dplyr::mutate(across(where(is.numeric), ~replace_na(.x, 0)))
data.test <- data.test %>% dplyr::mutate(across(where(is.numeric), ~replace_na(.x, 0)))

#set tasks for training and test datasets
task_train <- as_task_classif(x = data.train, target = 'pa', positive = '1')
task_test <- as_task_classif(x = data.test, target = 'pa', positive = '1')

# Make learner.
learner <-lrn(
  "classif.ranger",
  predict_type = "prob",
  mtry  = to_tune(1, ncol(data.train) - 4),
  sample.fraction = to_tune(0.2, 0.9),
  min.node.size = to_tune(p_int(1,10)),
  importance = 'impurity'
)

#tune hyperparameters
instance = ti(
  task = task_train,
  learner = learner,
  resampling = rsmp("cv", folds = 5),
  measures = msr("classif.ce"),
  terminator = trm("none")
)

tuner = mlr3tuning::tnr("grid_search", resolution = 5, batch_size = 5)
tuner$optimize(instance) # Takes ~ 4 minutes on my relatively fast computer

#store tuned hyperparameters in learner
learner$param_set$values <- instance$result_learner_param_vals

#finally! We can train our model with the train function
learner$train(task_train)

#let's quickly look at the model
learner$model

#Accuracy of model - first on training data, then testing data
measures <- msrs(c('classif.acc'))

pred_train <- learner$predict(task_train)
pred_train$confusion
pred_train$score(measures)


pred_test <- learner$predict(task_test)
pred_test$confusion
pred_test$score(measures)

#importance with iml package - this is looking at the most influencial predictors in the model
x_data <- data.train %>% dplyr::select(-pa) 
# Create "Predictor" object to interpret findings via the iml package.
predictor_data <- Predictor$new(learner, data = x_data, y = data.train$pa) 

imp_data <- FeatureImp$new(predictor_data, loss = "ce") # Calculate importance.

imp_df_data <- imp_data$results 

#also doesn't work because of the amount of 0's
```

```{r Binary_blue_crab}
#filter for blue crab
data <- st %>% filter(species_common == 'blue crab')

sum(data$pa)

data$pa <- as.factor(data$pa)

#Take out columns that will not be in model
data <- data %>% dplyr::select(-c(station, sample_period, species_common, species_abundance, secchi, pull_id, site))

str(data)
data$depth <- as.numeric(data$depth)
data$station_seagrass_richness <- as.numeric(data$station_seagrass_richness)
set.seed(19)
# Randomly select 70% of the data frame for the training dataset - from each
datasled <- data %>% dplyr::filter(type == "sled")
datatr <- data %>% dplyr::filter(type == 'trawl')
data.trains <- datasled[sample(1:nrow(datasled), nrow(datasled) * 0.8, replace = FALSE), ]
data.traint <- datatr[sample(1:nrow(datatr), nrow(datatr) * 0.8, replace = FALSE), ]
# Remainder (i.e., 30%) becomes the test dataset.
data.tests <- datasled[!(datasled$ID %in% data.trains$ID), ]
data.testt <- datatr[!(datatr$ID %in% data.traint$ID), ]


data.train <- bind_rows(data.trains, data.traint) %>% dplyr::select(-c(ID, tow_id))
data.test <- bind_rows(data.tests, data.testt) %>% dplyr::select(-c(ID, tow_id))

#remove columns with Inf data
data.train <- data.train %>% dplyr::select(where(~ !is.numeric(.x) || !any(is.infinite(.x))))
data.test <- data.test %>% dplyr::select(where(~ !is.numeric(.x) || !any(is.infinite(.x))))

#change columns with NA data to zeros
data.train <- data.train %>% dplyr::mutate(across(where(is.numeric), ~replace_na(.x, 0)))
data.test <- data.test %>% dplyr::mutate(across(where(is.numeric), ~replace_na(.x, 0)))

#set tasks for training and test datasets
task_train <- as_task_classif(x = data.train, target = 'pa', positive = '1')
task_test <- as_task_classif(x = data.test, target = 'pa', positive = '1')

# Make learner.
learner <-lrn(
  "classif.ranger",
  predict_type = "prob",
  mtry  = to_tune(1, ncol(data.train) - 4),
  sample.fraction = to_tune(0.2, 0.9),
  min.node.size = to_tune(p_int(1,10)),
  importance = 'impurity'
)

#tune hyperparameters
instance = ti(
  task = task_train,
  learner = learner,
  resampling = rsmp("cv", folds = 5),
  measures = msr("classif.ce"),
  terminator = trm("none")
)

tuner = mlr3tuning::tnr("grid_search", resolution = 5, batch_size = 5)
tuner$optimize(instance) # Takes ~ 4 minutes on my relatively fast computer

#store tuned hyperparameters in learner
learner$param_set$values <- instance$result_learner_param_vals

#finally! We can train our model with the train function
learner$train(task_train)

#let's quickly look at the model
learner$model

#Accuracy of model - first on training data, then testing data
measures <- msrs(c('classif.acc'))

pred_train <- learner$predict(task_train)
pred_train$confusion
pred_train$score(measures)


pred_test <- learner$predict(task_test)
pred_test$confusion
pred_test$score(measures)

#importance with iml package - this is looking at the most influencial predictors in the model
x_data <- data.train %>% dplyr::select(-pa) 
# Create "Predictor" object to interpret findings via the iml package.
predictor_data <- Predictor$new(learner, data = x_data, y = data.train$pa) 

imp_data <- FeatureImp$new(predictor_data, loss = "ce") # Calculate importance.

imp_df_data <- imp_data$results 

#very inaccurate
```

```{r Binary_pink_shrimp}
#filter for blue crab
data <- st %>% filter(species_common == 'pink shrimp')

sum(data$pa)

data$pa <- as.factor(data$pa)

#Take out columns that will not be in model
data <- data %>% dplyr::select(-c(station, sample_period, species_common, species_abundance, secchi, pull_id, site))

str(data)
data$depth <- as.numeric(data$depth)
data$station_seagrass_richness <- as.numeric(data$station_seagrass_richness)
set.seed(19)
# Randomly select 70% of the data frame for the training dataset - from each
datasled <- data %>% dplyr::filter(type == "sled")
datatr <- data %>% dplyr::filter(type == 'trawl')
data.trains <- datasled[sample(1:nrow(datasled), nrow(datasled) * 0.8, replace = FALSE), ]
data.traint <- datatr[sample(1:nrow(datatr), nrow(datatr) * 0.8, replace = FALSE), ]
# Remainder (i.e., 30%) becomes the test dataset.
data.tests <- datasled[!(datasled$ID %in% data.trains$ID), ]
data.testt <- datatr[!(datatr$ID %in% data.traint$ID), ]


data.train <- bind_rows(data.trains, data.traint) %>% dplyr::select(-c(ID, tow_id))
data.test <- bind_rows(data.tests, data.testt) %>% dplyr::select(-c(ID, tow_id))

#remove columns with Inf data
data.train <- data.train %>% dplyr::select(where(~ !is.numeric(.x) || !any(is.infinite(.x))))
data.test <- data.test %>% dplyr::select(where(~ !is.numeric(.x) || !any(is.infinite(.x))))

#change columns with NA data to zeros
data.train <- data.train %>% dplyr::mutate(across(where(is.numeric), ~replace_na(.x, 0)))
data.test <- data.test %>% dplyr::mutate(across(where(is.numeric), ~replace_na(.x, 0)))

#set tasks for training and test datasets
task_train <- as_task_classif(x = data.train, target = 'pa', positive = '1')
task_test <- as_task_classif(x = data.test, target = 'pa', positive = '1')

# Make learner.
learner <-lrn(
  "classif.ranger",
  predict_type = "prob",
  mtry  = to_tune(1, ncol(data.train) - 4),
  sample.fraction = to_tune(0.2, 0.9),
  min.node.size = to_tune(p_int(1,10)),
  importance = 'impurity'
)

#tune hyperparameters
instance = ti(
  task = task_train,
  learner = learner,
  resampling = rsmp("cv", folds = 5),
  measures = msr("classif.ce"),
  terminator = trm("none")
)

tuner = mlr3tuning::tnr("grid_search", resolution = 5, batch_size = 5)
tuner$optimize(instance) # Takes ~ 4 minutes on my relatively fast computer

#store tuned hyperparameters in learner
learner$param_set$values <- instance$result_learner_param_vals

#finally! We can train our model with the train function
learner$train(task_train)

#let's quickly look at the model
learner$model

#Accuracy of model - first on training data, then testing data
measures <- msrs(c('classif.acc'))

pred_train <- learner$predict(task_train)
pred_train$confusion
pred_train$score(measures)


pred_test <- learner$predict(task_test)
pred_test$confusion
pred_test$score(measures)

#importance with iml package - this is looking at the most influencial predictors in the model
x_data <- data.train %>% dplyr::select(-pa) 
# Create "Predictor" object to interpret findings via the iml package.
predictor_data <- Predictor$new(learner, data = x_data, y = data.train$pa) 

imp_data <- FeatureImp$new(predictor_data, loss = "ce") # Calculate importance.

imp_df_data <- imp_data$results 

#not too bad in accuracy, but still not great. Salinity is the most important factor
effect_sal <- FeatureEffect$new(predictor_data, feature = c('salinity'), method = 'pdp')
effect_sal$plot()
salpinkshrimp <- effect_sal$plot()

#The next important one is se of canopy height
effect_chse <- FeatureEffect$new(predictor_data, feature = c('station_se_mcanopy'), method = 'pdp')
effect_chse$plot()
chsepinkshrimp <- effect_chse$plot()

#looks like pink shrimp are more present in lower salinities, but have the same relationship with canopy height as brown shrimp
```

```{r Binary_gray_snapper}
#filter for gray snapper
data <- st %>% filter(species_common == 'gray snapper')

sum(data$pa)

data$pa <- as.factor(data$pa)

#Take out columns that will not be in model
data <- data %>% dplyr::select(-c(station, sample_period, species_common, species_abundance, secchi, pull_id, site))

str(data)
data$depth <- as.numeric(data$depth)
data$station_seagrass_richness <- as.numeric(data$station_seagrass_richness)
set.seed(19)
# Randomly select 70% of the data frame for the training dataset - from each
datasled <- data %>% dplyr::filter(type == "sled")
datatr <- data %>% dplyr::filter(type == 'trawl')
data.trains <- datasled[sample(1:nrow(datasled), nrow(datasled) * 0.8, replace = FALSE), ]
data.traint <- datatr[sample(1:nrow(datatr), nrow(datatr) * 0.8, replace = FALSE), ]
# Remainder (i.e., 30%) becomes the test dataset.
data.tests <- datasled[!(datasled$ID %in% data.trains$ID), ]
data.testt <- datatr[!(datatr$ID %in% data.traint$ID), ]


data.train <- bind_rows(data.trains, data.traint) %>% dplyr::select(-c(ID, tow_id))
data.test <- bind_rows(data.tests, data.testt) %>% dplyr::select(-c(ID, tow_id))

#remove columns with Inf data
data.train <- data.train %>% dplyr::select(where(~ !is.numeric(.x) || !any(is.infinite(.x))))
data.test <- data.test %>% dplyr::select(where(~ !is.numeric(.x) || !any(is.infinite(.x))))

#change columns with NA data to zeros
data.train <- data.train %>% dplyr::mutate(across(where(is.numeric), ~replace_na(.x, 0)))
data.test <- data.test %>% dplyr::mutate(across(where(is.numeric), ~replace_na(.x, 0)))

#set tasks for training and test datasets
task_train <- as_task_classif(x = data.train, target = 'pa', positive = '1')
task_test <- as_task_classif(x = data.test, target = 'pa', positive = '1')

# Make learner.
learner <-lrn(
  "classif.ranger",
  predict_type = "prob",
  mtry  = to_tune(1, ncol(data.train) - 4),
  sample.fraction = to_tune(0.2, 0.9),
  min.node.size = to_tune(p_int(1,10)),
  importance = 'impurity'
)

#tune hyperparameters
instance = ti(
  task = task_train,
  learner = learner,
  resampling = rsmp("cv", folds = 5),
  measures = msr("classif.ce"),
  terminator = trm("none")
)

tuner = mlr3tuning::tnr("grid_search", resolution = 5, batch_size = 5)
tuner$optimize(instance) # Takes ~ 4 minutes on my relatively fast computer

#store tuned hyperparameters in learner
learner$param_set$values <- instance$result_learner_param_vals

#finally! We can train our model with the train function
learner$train(task_train)

#let's quickly look at the model
learner$model

#Accuracy of model - first on training data, then testing data
measures <- msrs(c('classif.acc'))

pred_train <- learner$predict(task_train)
pred_train$confusion
pred_train$score(measures)


pred_test <- learner$predict(task_test)
pred_test$confusion
pred_test$score(measures)

#importance with iml package - this is looking at the most influencial predictors in the model
x_data <- data.train %>% dplyr::select(-pa) 
# Create "Predictor" object to interpret findings via the iml package.
predictor_data <- Predictor$new(learner, data = x_data, y = data.train$pa) 

imp_data <- FeatureImp$new(predictor_data, loss = "ce") # Calculate importance.

imp_df_data <- imp_data$results 

#awful accuracy!
```

#The binary models with all of the parameters were not very accurate, with the brown shrimp model the only one I trust. Let's try reducing the number of parameters to see if it improves the model

```{r Binary_subset_brown_shrimp}
#filter for brown shrimp
data <- st %>% filter(species_common == 'brown shrimp')

data$pa <- as.factor(data$pa)

#Take out columns that will not be in model
data <- data %>% dplyr::select(-c(station, sample_period, species_common, species_abundance, secchi, pull_id, site))

str(data)
data$depth <- as.numeric(data$depth)
data$station_seagrass_richness <- as.numeric(data$station_seagrass_richness)

#subset parameters
data <- data %>% dplyr::select(station_total_pcov, station_se_total_pcov, station_tt_pcov, station_hw_pcov, station_sf_pcov, station_mac_pcov, station_drift_pcov, depth, temp, salinity, pa, seagrass_richness, type, do, light_attenuation_av, station_leaf_mlength, station_se_mcanopy, station_shoots_m, station_se_shoots_m, station_ep_wt_gg, ID)
#this subset covers all of the water quality parameters, total percent cover of the SAV as well as the individual species, seagrass richness, canopy height, density of shoots, and epiphytic load

# Randomly select 70% of the data frame for the training dataset - from each
set.seed(19)
datasled <- data %>% dplyr::filter(type == "sled")
datatr <- data %>% dplyr::filter(type == 'trawl')
data.trains <- datasled[sample(1:nrow(datasled), nrow(datasled) * 0.8, replace = FALSE), ]
data.traint <- datatr[sample(1:nrow(datatr), nrow(datatr) * 0.8, replace = FALSE), ]
# Remainder (i.e., 30%) becomes the test dataset.
data.tests <- datasled[!(datasled$ID %in% data.trains$ID), ]
data.testt <- datatr[!(datatr$ID %in% data.traint$ID), ]


data.train <- bind_rows(data.trains, data.traint) %>% dplyr::select(-c(ID))
data.test <- bind_rows(data.tests, data.testt) %>% dplyr::select(-c(ID))

#remove columns with Inf data
data.train <- data.train %>% dplyr::select(where(~ !is.numeric(.x) || !any(is.infinite(.x))))
data.test <- data.test %>% dplyr::select(where(~ !is.numeric(.x) || !any(is.infinite(.x))))

#change columns with NA data to zeros
data.train <- data.train %>% dplyr::mutate(across(where(is.numeric), ~replace_na(.x, 0)))
data.test <- data.test %>% dplyr::mutate(across(where(is.numeric), ~replace_na(.x, 0)))


#set tasks for training and test datasets
task_train <- as_task_classif(x = data.train, target = 'pa', positive = '1')
task_test <- as_task_classif(x = data.test, target = 'pa', positive = '1')

# Make learner.
learner <-lrn(
  "classif.ranger",
  predict_type = "prob",
  mtry  = to_tune(1, ncol(data.train) - 4),
  sample.fraction = to_tune(0.2, 0.9),
  min.node.size = to_tune(p_int(1,10)),
  importance = 'impurity'
)

#tune hyperparameters
instance = ti(
  task = task_train,
  learner = learner,
  resampling = rsmp("cv", folds = 5),
  measures = msr("classif.ce"),
  terminator = trm("none")
)

tuner = mlr3tuning::tnr("grid_search", resolution = 5, batch_size = 5)
tuner$optimize(instance) # Takes ~ 4 minutes on my relatively fast computer

#store tuned hyperparameters in learner
learner$param_set$values <- instance$result_learner_param_vals

#finally! We can train our model with the train function
learner$train(task_train)

#let's quickly look at the model
learner$model

#Accuracy of model - first on training data, then testing data
measures <- msrs(c('classif.acc'))

pred_train <- learner$predict(task_train)
pred_train$confusion
pred_train$score(measures)


pred_test <- learner$predict(task_test)
pred_test$confusion
pred_test$score(measures)

#importance with iml package - this is looking at the most influencial predictors in the model
x_data <- data.train %>% dplyr::select(-pa) 
# Create "Predictor" object to interpret findings via the iml package.
predictor_data <- Predictor$new(learner, data = x_data, y = data.train$pa) 

imp_data <- FeatureImp$new(predictor_data, loss = "ce") # Calculate importance.

imp_df_data <- imp_data$results 

#so salinity is the most important variable for brown shrimp out of the environmental data. Let's investigate how it impacts the presence of shrimp

effect_sal <- FeatureEffect$new(predictor_data, feature = c('salinity'), method = 'pdp')
effect_sal$plot()
salbrownshrimp_subset <- effect_sal$plot()

#most important covariate is the type of collection. Here is what it looks like
effect_type <- FeatureEffect$new(predictor_data, feature = c('type'), method = 'pdp')
effect_type$plot()
typebrownshrimp <- effect_type$plot()

#The next is actually epiphytic load!

effect_epi <- FeatureEffect$new(predictor_data, feature = c('station_ep_wt_gg'), method = 'pdp')
effect_epi$plot()
epibrownshrimp_subset <- effect_epi$plot()

#Overall, lower probability of presence at medium salinities and lower amount of epiphytes. More likely to catch brown shrimp in sled
```

```{r Binary_subset_pink_shrimp}
#filter for pink shrimp
data <- st %>% filter(species_common == 'pink shrimp')

data$pa <- as.factor(data$pa)

#Take out columns that will not be in model
data <- data %>% dplyr::select(-c(station, sample_period, species_common, species_abundance, secchi, pull_id, site))

str(data)
data$depth <- as.numeric(data$depth)
data$station_seagrass_richness <- as.numeric(data$station_seagrass_richness)

#subset parameters
data <- data %>% dplyr::select(station_total_pcov, station_se_total_pcov, station_tt_pcov, station_hw_pcov, station_sf_pcov, station_mac_pcov, station_drift_pcov, depth, temp, salinity, pa, seagrass_richness, type, do, light_attenuation_av, station_leaf_mlength, station_se_mcanopy, station_shoots_m, station_se_shoots_m, station_ep_wt_gg, ID)
#this subset covers all of the water quality parameters, total percent cover of the SAV as well as the individual species, seagrass richness, canopy height, density of shoots, and epiphytic load

# Randomly select 70% of the data frame for the training dataset - from each
set.seed(19)
datasled <- data %>% dplyr::filter(type == "sled")
datatr <- data %>% dplyr::filter(type == 'trawl')
data.trains <- datasled[sample(1:nrow(datasled), nrow(datasled) * 0.8, replace = FALSE), ]
data.traint <- datatr[sample(1:nrow(datatr), nrow(datatr) * 0.8, replace = FALSE), ]
# Remainder (i.e., 30%) becomes the test dataset.
data.tests <- datasled[!(datasled$ID %in% data.trains$ID), ]
data.testt <- datatr[!(datatr$ID %in% data.traint$ID), ]


data.train <- bind_rows(data.trains, data.traint) %>% dplyr::select(-c(ID))
data.test <- bind_rows(data.tests, data.testt) %>% dplyr::select(-c(ID))

#remove columns with Inf data
data.train <- data.train %>% dplyr::select(where(~ !is.numeric(.x) || !any(is.infinite(.x))))
data.test <- data.test %>% dplyr::select(where(~ !is.numeric(.x) || !any(is.infinite(.x))))

#change columns with NA data to zeros
data.train <- data.train %>% dplyr::mutate(across(where(is.numeric), ~replace_na(.x, 0)))
data.test <- data.test %>% dplyr::mutate(across(where(is.numeric), ~replace_na(.x, 0)))


#set tasks for training and test datasets
task_train <- as_task_classif(x = data.train, target = 'pa', positive = '1')
task_test <- as_task_classif(x = data.test, target = 'pa', positive = '1')

# Make learner.
learner <-lrn(
  "classif.ranger",
  predict_type = "prob",
  mtry  = to_tune(1, ncol(data.train) - 4),
  sample.fraction = to_tune(0.2, 0.9),
  min.node.size = to_tune(p_int(1,10)),
  importance = 'impurity'
)

#tune hyperparameters
instance = ti(
  task = task_train,
  learner = learner,
  resampling = rsmp("cv", folds = 5),
  measures = msr("classif.ce"),
  terminator = trm("none")
)

tuner = mlr3tuning::tnr("grid_search", resolution = 5, batch_size = 5)
tuner$optimize(instance) # Takes ~ 4 minutes on my relatively fast computer

#store tuned hyperparameters in learner
learner$param_set$values <- instance$result_learner_param_vals

#finally! We can train our model with the train function
learner$train(task_train)

#let's quickly look at the model
learner$model

#Accuracy of model - first on training data, then testing data
measures <- msrs(c('classif.acc'))

pred_train <- learner$predict(task_train)
pred_train$confusion
pred_train$score(measures)


pred_test <- learner$predict(task_test)
pred_test$confusion
pred_test$score(measures)

#importance with iml package - this is looking at the most influencial predictors in the model
x_data <- data.train %>% dplyr::select(-pa) 
# Create "Predictor" object to interpret findings via the iml package.
predictor_data <- Predictor$new(learner, data = x_data, y = data.train$pa) 

imp_data <- FeatureImp$new(predictor_data, loss = "ce") # Calculate importance.

imp_df_data <- imp_data$results 

#so salinity is the most important variable for pink shrimp out of the environmental data. Let's investigate how it impacts the presence of shrimp

effect_sal <- FeatureEffect$new(predictor_data, feature = c('salinity'), method = 'pdp')
effect_sal$plot()
salpinkshrimp_subset <- effect_sal$plot()

#Not super accurate, but decent again
```

```{r Binary_subset_blue_crab}
#filter for blue crab
data <- st %>% filter(species_common == 'blue crab')

data$pa <- as.factor(data$pa)

#Take out columns that will not be in model
data <- data %>% dplyr::select(-c(station, sample_period, species_common, species_abundance, secchi, pull_id, site))

str(data)
data$depth <- as.numeric(data$depth)
data$station_seagrass_richness <- as.numeric(data$station_seagrass_richness)

#subset parameters
data <- data %>% dplyr::select(station_total_pcov, station_se_total_pcov, station_tt_pcov, station_hw_pcov, station_sf_pcov, station_mac_pcov, station_drift_pcov, depth, temp, salinity, pa, seagrass_richness, type, do, light_attenuation_av, station_leaf_mlength, station_se_mcanopy, station_shoots_m, station_se_shoots_m, station_ep_wt_gg, ID)
#this subset covers all of the water quality parameters, total percent cover of the SAV as well as the individual species, seagrass richness, canopy height, density of shoots, and epiphytic load

# Randomly select 70% of the data frame for the training dataset - from each
set.seed(19)
datasled <- data %>% dplyr::filter(type == "sled")
datatr <- data %>% dplyr::filter(type == 'trawl')
data.trains <- datasled[sample(1:nrow(datasled), nrow(datasled) * 0.8, replace = FALSE), ]
data.traint <- datatr[sample(1:nrow(datatr), nrow(datatr) * 0.8, replace = FALSE), ]
# Remainder (i.e., 30%) becomes the test dataset.
data.tests <- datasled[!(datasled$ID %in% data.trains$ID), ]
data.testt <- datatr[!(datatr$ID %in% data.traint$ID), ]


data.train <- bind_rows(data.trains, data.traint) %>% dplyr::select(-c(ID))
data.test <- bind_rows(data.tests, data.testt) %>% dplyr::select(-c(ID))

#remove columns with Inf data
data.train <- data.train %>% dplyr::select(where(~ !is.numeric(.x) || !any(is.infinite(.x))))
data.test <- data.test %>% dplyr::select(where(~ !is.numeric(.x) || !any(is.infinite(.x))))

#change columns with NA data to zeros
data.train <- data.train %>% dplyr::mutate(across(where(is.numeric), ~replace_na(.x, 0)))
data.test <- data.test %>% dplyr::mutate(across(where(is.numeric), ~replace_na(.x, 0)))


#set tasks for training and test datasets
task_train <- as_task_classif(x = data.train, target = 'pa', positive = '1')
task_test <- as_task_classif(x = data.test, target = 'pa', positive = '1')

# Make learner.
learner <-lrn(
  "classif.ranger",
  predict_type = "prob",
  mtry  = to_tune(1, ncol(data.train) - 4),
  sample.fraction = to_tune(0.2, 0.9),
  min.node.size = to_tune(p_int(1,10)),
  importance = 'impurity'
)

#tune hyperparameters
instance = ti(
  task = task_train,
  learner = learner,
  resampling = rsmp("cv", folds = 5),
  measures = msr("classif.ce"),
  terminator = trm("none")
)

tuner = mlr3tuning::tnr("grid_search", resolution = 5, batch_size = 5)
tuner$optimize(instance) # Takes ~ 4 minutes on my relatively fast computer

#store tuned hyperparameters in learner
learner$param_set$values <- instance$result_learner_param_vals

#finally! We can train our model with the train function
learner$train(task_train)

#let's quickly look at the model
learner$model

#Accuracy of model - first on training data, then testing data
measures <- msrs(c('classif.acc'))

pred_train <- learner$predict(task_train)
pred_train$confusion
pred_train$score(measures)


pred_test <- learner$predict(task_test)
pred_test$confusion
pred_test$score(measures)

#importance with iml package - this is looking at the most influencial predictors in the model
x_data <- data.train %>% dplyr::select(-pa) 
# Create "Predictor" object to interpret findings via the iml package.
predictor_data <- Predictor$new(learner, data = x_data, y = data.train$pa) 

imp_data <- FeatureImp$new(predictor_data, loss = "ce") # Calculate importance.

imp_df_data <- imp_data$results 

#very inaccurate
```

```{r Binary_subset_spotted_seatrout}
#filter for pink shrimp
data <- st %>% filter(species_common == 'spotted seatrout')

data$pa <- as.factor(data$pa)

#Take out columns that will not be in model
data <- data %>% dplyr::select(-c(station, sample_period, species_common, species_abundance, secchi, pull_id, site))

str(data)
data$depth <- as.numeric(data$depth)
data$station_seagrass_richness <- as.numeric(data$station_seagrass_richness)

#subset parameters
data <- data %>% dplyr::select(station_total_pcov, station_se_total_pcov, station_tt_pcov, station_hw_pcov, station_sf_pcov, station_mac_pcov, station_drift_pcov, depth, temp, salinity, pa, seagrass_richness, type, do, light_attenuation_av, station_leaf_mlength, station_se_mcanopy, station_shoots_m, station_se_shoots_m, station_ep_wt_gg, ID)
#this subset covers all of the water quality parameters, total percent cover of the SAV as well as the individual species, seagrass richness, canopy height, density of shoots, and epiphytic load

# Randomly select 70% of the data frame for the training dataset - from each
set.seed(19)
datasled <- data %>% dplyr::filter(type == "sled")
datatr <- data %>% dplyr::filter(type == 'trawl')
data.trains <- datasled[sample(1:nrow(datasled), nrow(datasled) * 0.8, replace = FALSE), ]
data.traint <- datatr[sample(1:nrow(datatr), nrow(datatr) * 0.8, replace = FALSE), ]
# Remainder (i.e., 30%) becomes the test dataset.
data.tests <- datasled[!(datasled$ID %in% data.trains$ID), ]
data.testt <- datatr[!(datatr$ID %in% data.traint$ID), ]


data.train <- bind_rows(data.trains, data.traint) %>% dplyr::select(-c(ID))
data.test <- bind_rows(data.tests, data.testt) %>% dplyr::select(-c(ID))

#remove columns with Inf data
data.train <- data.train %>% dplyr::select(where(~ !is.numeric(.x) || !any(is.infinite(.x))))
data.test <- data.test %>% dplyr::select(where(~ !is.numeric(.x) || !any(is.infinite(.x))))

#change columns with NA data to zeros
data.train <- data.train %>% dplyr::mutate(across(where(is.numeric), ~replace_na(.x, 0)))
data.test <- data.test %>% dplyr::mutate(across(where(is.numeric), ~replace_na(.x, 0)))


#set tasks for training and test datasets
task_train <- as_task_classif(x = data.train, target = 'pa', positive = '1')
task_test <- as_task_classif(x = data.test, target = 'pa', positive = '1')

# Make learner.
learner <-lrn(
  "classif.ranger",
  predict_type = "prob",
  mtry  = to_tune(1, ncol(data.train) - 4),
  sample.fraction = to_tune(0.2, 0.9),
  min.node.size = to_tune(p_int(1,10)),
  importance = 'impurity'
)

#tune hyperparameters
instance = ti(
  task = task_train,
  learner = learner,
  resampling = rsmp("cv", folds = 5),
  measures = msr("classif.ce"),
  terminator = trm("none")
)

tuner = mlr3tuning::tnr("grid_search", resolution = 5, batch_size = 5)
tuner$optimize(instance) # Takes ~ 4 minutes on my relatively fast computer

#store tuned hyperparameters in learner
learner$param_set$values <- instance$result_learner_param_vals

#finally! We can train our model with the train function
learner$train(task_train)

#let's quickly look at the model
learner$model

#Accuracy of model - first on training data, then testing data
measures <- msrs(c('classif.acc'))

pred_train <- learner$predict(task_train)
pred_train$confusion
pred_train$score(measures)


pred_test <- learner$predict(task_test)
pred_test$confusion
pred_test$score(measures)

#All but one data point was predicted to be a zero. Nope
```

```{r Binary_subset_gray_snapper}
#filter for gray snapper
data <- st %>% filter(species_common == 'gray snapper')

data$pa <- as.factor(data$pa)

#Take out columns that will not be in model
data <- data %>% dplyr::select(-c(station, sample_period, species_common, species_abundance, secchi, pull_id, site))

str(data)
data$depth <- as.numeric(data$depth)
data$station_seagrass_richness <- as.numeric(data$station_seagrass_richness)

#subset parameters
data <- data %>% dplyr::select(station_total_pcov, station_se_total_pcov, station_tt_pcov, station_hw_pcov, station_sf_pcov, station_mac_pcov, station_drift_pcov, depth, temp, salinity, pa, seagrass_richness, type, do, light_attenuation_av, station_leaf_mlength, station_se_mcanopy, station_shoots_m, station_se_shoots_m, station_ep_wt_gg, ID)
#this subset covers all of the water quality parameters, total percent cover of the SAV as well as the individual species, seagrass richness, canopy height, density of shoots, and epiphytic load

# Randomly select 70% of the data frame for the training dataset - from each
set.seed(19)
datasled <- data %>% dplyr::filter(type == "sled")
datatr <- data %>% dplyr::filter(type == 'trawl')
data.trains <- datasled[sample(1:nrow(datasled), nrow(datasled) * 0.8, replace = FALSE), ]
data.traint <- datatr[sample(1:nrow(datatr), nrow(datatr) * 0.8, replace = FALSE), ]
# Remainder (i.e., 30%) becomes the test dataset.
data.tests <- datasled[!(datasled$ID %in% data.trains$ID), ]
data.testt <- datatr[!(datatr$ID %in% data.traint$ID), ]


data.train <- bind_rows(data.trains, data.traint) %>% dplyr::select(-c(ID))
data.test <- bind_rows(data.tests, data.testt) %>% dplyr::select(-c(ID))

#remove columns with Inf data
data.train <- data.train %>% dplyr::select(where(~ !is.numeric(.x) || !any(is.infinite(.x))))
data.test <- data.test %>% dplyr::select(where(~ !is.numeric(.x) || !any(is.infinite(.x))))

#change columns with NA data to zeros
data.train <- data.train %>% dplyr::mutate(across(where(is.numeric), ~replace_na(.x, 0)))
data.test <- data.test %>% dplyr::mutate(across(where(is.numeric), ~replace_na(.x, 0)))


#set tasks for training and test datasets
task_train <- as_task_classif(x = data.train, target = 'pa', positive = '1')
task_test <- as_task_classif(x = data.test, target = 'pa', positive = '1')

# Make learner.
learner <-lrn(
  "classif.ranger",
  predict_type = "prob",
  mtry  = to_tune(1, ncol(data.train) - 4),
  sample.fraction = to_tune(0.2, 0.9),
  min.node.size = to_tune(p_int(1,10)),
  importance = 'impurity'
)

#tune hyperparameters
instance = ti(
  task = task_train,
  learner = learner,
  resampling = rsmp("cv", folds = 5),
  measures = msr("classif.ce"),
  terminator = trm("none")
)

tuner = mlr3tuning::tnr("grid_search", resolution = 5, batch_size = 5)
tuner$optimize(instance) # Takes ~ 4 minutes on my relatively fast computer

#store tuned hyperparameters in learner
learner$param_set$values <- instance$result_learner_param_vals

#finally! We can train our model with the train function
learner$train(task_train)

#let's quickly look at the model
learner$model

#Accuracy of model - first on training data, then testing data
measures <- msrs(c('classif.acc'))

pred_train <- learner$predict(task_train)
pred_train$confusion
pred_train$score(measures)


pred_test <- learner$predict(task_test)
pred_test$confusion
pred_test$score(measures)

#importance with iml package - this is looking at the most influencial predictors in the model
x_data <- data.train %>% dplyr::select(-pa) 
# Create "Predictor" object to interpret findings via the iml package.
predictor_data <- Predictor$new(learner, data = x_data, y = data.train$pa) 

imp_data <- FeatureImp$new(predictor_data, loss = "ce") # Calculate importance.

imp_df_data <- imp_data$results 

#gray snapper model actually worked! Not great, but it is better. Best predictor is gear type
effect_type <- FeatureEffect$new(predictor_data, feature = c('type'), method = 'pdp')
effect_type$plot()
typegraysnapper_subset <- effect_type$plot()

#next were 2 habitat things, epiphytic load and cover of thalassia. Then it is temperature and salinity
effect_epi <- FeatureEffect$new(predictor_data, feature = c('station_ep_wt_gg'), method = 'pdp')
effect_epi$plot()
epigraysnapper_subset <- effect_epi$plot()

effect_tt <- FeatureEffect$new(predictor_data, feature = c('station_tt_pcov'), method = 'pdp')
effect_tt$plot()
ttgraysnapper_subset <- effect_tt$plot()

effect_temp <- FeatureEffect$new(predictor_data, feature = c('temp'), method = 'pdp')
effect_temp$plot()
tempgraysnapper_subset <- effect_temp$plot()

effect_sal <- FeatureEffect$new(predictor_data, feature = c('salinity'), method = 'pdp')
effect_sal$plot()
salgraysnapper_subset <- effect_sal$plot()

#Not great, but seems like snapper are more likely to be caught in trawls, found in areas with higher epiphytic biomass and higher thalassia cover. I wouldn't trust temp/sal data
```

```{r Binary_subset_lane_snapper}
#filter for lane snapper
data <- st %>% filter(species_common == 'lane snapper')

data$pa <- as.factor(data$pa)

#Take out columns that will not be in model
data <- data %>% dplyr::select(-c(station, sample_period, species_common, species_abundance, secchi, pull_id, site))

str(data)
data$depth <- as.numeric(data$depth)
data$station_seagrass_richness <- as.numeric(data$station_seagrass_richness)

#subset parameters
data <- data %>% dplyr::select(station_total_pcov, station_se_total_pcov, station_tt_pcov, station_hw_pcov, station_sf_pcov, station_mac_pcov, station_drift_pcov, depth, temp, salinity, pa, seagrass_richness, type, do, light_attenuation_av, station_leaf_mlength, station_se_mcanopy, station_shoots_m, station_se_shoots_m, station_ep_wt_gg, ID)
#this subset covers all of the water quality parameters, total percent cover of the SAV as well as the individual species, seagrass richness, canopy height, density of shoots, and epiphytic load

# Randomly select 70% of the data frame for the training dataset - from each
set.seed(19)
datasled <- data %>% dplyr::filter(type == "sled")
datatr <- data %>% dplyr::filter(type == 'trawl')
data.trains <- datasled[sample(1:nrow(datasled), nrow(datasled) * 0.8, replace = FALSE), ]
data.traint <- datatr[sample(1:nrow(datatr), nrow(datatr) * 0.8, replace = FALSE), ]
# Remainder (i.e., 30%) becomes the test dataset.
data.tests <- datasled[!(datasled$ID %in% data.trains$ID), ]
data.testt <- datatr[!(datatr$ID %in% data.traint$ID), ]


data.train <- bind_rows(data.trains, data.traint) %>% dplyr::select(-c(ID))
data.test <- bind_rows(data.tests, data.testt) %>% dplyr::select(-c(ID))

#remove columns with Inf data
data.train <- data.train %>% dplyr::select(where(~ !is.numeric(.x) || !any(is.infinite(.x))))
data.test <- data.test %>% dplyr::select(where(~ !is.numeric(.x) || !any(is.infinite(.x))))

#change columns with NA data to zeros
data.train <- data.train %>% dplyr::mutate(across(where(is.numeric), ~replace_na(.x, 0)))
data.test <- data.test %>% dplyr::mutate(across(where(is.numeric), ~replace_na(.x, 0)))


#set tasks for training and test datasets
task_train <- as_task_classif(x = data.train, target = 'pa', positive = '1')
task_test <- as_task_classif(x = data.test, target = 'pa', positive = '1')

# Make learner.
learner <-lrn(
  "classif.ranger",
  predict_type = "prob",
  mtry  = to_tune(1, ncol(data.train) - 4),
  sample.fraction = to_tune(0.2, 0.9),
  min.node.size = to_tune(p_int(1,10)),
  importance = 'impurity'
)

#tune hyperparameters
instance = ti(
  task = task_train,
  learner = learner,
  resampling = rsmp("cv", folds = 5),
  measures = msr("classif.ce"),
  terminator = trm("none")
)

tuner = mlr3tuning::tnr("grid_search", resolution = 5, batch_size = 5)
tuner$optimize(instance) # Takes ~ 4 minutes on my relatively fast computer

#store tuned hyperparameters in learner
learner$param_set$values <- instance$result_learner_param_vals

#finally! We can train our model with the train function
learner$train(task_train)

#let's quickly look at the model
learner$model

#Accuracy of model - first on training data, then testing data
measures <- msrs(c('classif.acc'))

pred_train <- learner$predict(task_train)
pred_train$confusion
pred_train$score(measures)


pred_test <- learner$predict(task_test)
pred_test$confusion
pred_test$score(measures)

#importance with iml package - this is looking at the most influencial predictors in the model
x_data <- data.train %>% dplyr::select(-pa) 
# Create "Predictor" object to interpret findings via the iml package.
predictor_data <- Predictor$new(learner, data = x_data, y = data.train$pa) 

imp_data <- FeatureImp$new(predictor_data, loss = "ce") # Calculate importance.

imp_df_data <- imp_data$results 

#Lane snapper model not good
```

```{r Binary_subset_red_drum}
#filter for red drum
data <- st %>% filter(species_common == 'red drum')

sum(data$pa)

#only 5 samples had red drum. Can't run this

data$pa <- as.factor(data$pa)

#Take out columns that will not be in model
data <- data %>% dplyr::select(-c(station, sample_period, species_common, species_abundance, secchi, pull_id, site))

str(data)
data$depth <- as.numeric(data$depth)
data$station_seagrass_richness <- as.numeric(data$station_seagrass_richness)

#subset parameters
data <- data %>% dplyr::select(station_total_pcov, station_se_total_pcov, station_tt_pcov, station_hw_pcov, station_sf_pcov, station_mac_pcov, station_drift_pcov, depth, temp, salinity, pa, seagrass_richness, type, do, light_attenuation_av, station_leaf_mlength, station_se_mcanopy, station_shoots_m, station_se_shoots_m, station_ep_wt_gg, ID)
#this subset covers all of the water quality parameters, total percent cover of the SAV as well as the individual species, seagrass richness, canopy height, density of shoots, and epiphytic load

# Randomly select 70% of the data frame for the training dataset - from each
set.seed(19)
datasled <- data %>% dplyr::filter(type == "sled")
datatr <- data %>% dplyr::filter(type == 'trawl')
data.trains <- datasled[sample(1:nrow(datasled), nrow(datasled) * 0.8, replace = FALSE), ]
data.traint <- datatr[sample(1:nrow(datatr), nrow(datatr) * 0.8, replace = FALSE), ]
# Remainder (i.e., 30%) becomes the test dataset.
data.tests <- datasled[!(datasled$ID %in% data.trains$ID), ]
data.testt <- datatr[!(datatr$ID %in% data.traint$ID), ]


data.train <- bind_rows(data.trains, data.traint) %>% dplyr::select(-c(ID))
data.test <- bind_rows(data.tests, data.testt) %>% dplyr::select(-c(ID))

#remove columns with Inf data
data.train <- data.train %>% dplyr::select(where(~ !is.numeric(.x) || !any(is.infinite(.x))))
data.test <- data.test %>% dplyr::select(where(~ !is.numeric(.x) || !any(is.infinite(.x))))

#change columns with NA data to zeros
data.train <- data.train %>% dplyr::mutate(across(where(is.numeric), ~replace_na(.x, 0)))
data.test <- data.test %>% dplyr::mutate(across(where(is.numeric), ~replace_na(.x, 0)))


#set tasks for training and test datasets
task_train <- as_task_classif(x = data.train, target = 'pa', positive = '1')
task_test <- as_task_classif(x = data.test, target = 'pa', positive = '1')

# Make learner.
learner <-lrn(
  "classif.ranger",
  predict_type = "prob",
  mtry  = to_tune(1, ncol(data.train) - 4),
  sample.fraction = to_tune(0.2, 0.9),
  min.node.size = to_tune(p_int(1,10)),
  importance = 'impurity'
)

#tune hyperparameters
instance = ti(
  task = task_train,
  learner = learner,
  resampling = rsmp("cv", folds = 5),
  measures = msr("classif.ce"),
  terminator = trm("none")
)

tuner = mlr3tuning::tnr("grid_search", resolution = 5, batch_size = 5)
tuner$optimize(instance) # Takes ~ 4 minutes on my relatively fast computer

#store tuned hyperparameters in learner
learner$param_set$values <- instance$result_learner_param_vals

#finally! We can train our model with the train function
learner$train(task_train)

#let's quickly look at the model
learner$model

#Accuracy of model - first on training data, then testing data
measures <- msrs(c('classif.acc'))

pred_train <- learner$predict(task_train)
pred_train$confusion
pred_train$score(measures)


pred_test <- learner$predict(task_test)
pred_test$confusion
pred_test$score(measures)

#importance with iml package - this is looking at the most influencial predictors in the model
x_data <- data.train %>% dplyr::select(-pa) 
# Create "Predictor" object to interpret findings via the iml package.
predictor_data <- Predictor$new(learner, data = x_data, y = data.train$pa) 

imp_data <- FeatureImp$new(predictor_data, loss = "ce") # Calculate importance.

imp_df_data <- imp_data$results 

#Lane snapper model not good
```