---
title: "RF_Nekton_Habitat_Suitability"
output: html_document
date: "2026-04-01"
---
## Author: Jon Rodemann

```{r setup, include=FALSE}
library(tidyverse)
library(randomForest)
library(mlr3verse)
library(iml)
library(ranger)
```

The following code goes through various iterations of random forest models to look at habitat suitability of nekton in seagrass habitats

```{r, include=FALSE}
#Load in datasets
sled <- read.csv('Data/combined_sled_species_sp.csv')
trawl <- read.csv('Data/combined_trawl_species.csv')

##We want to isolate the 8 important fish species. Here are the ones:
# - Gray snapper
# - Blue crab
# - Pink shrimp
# - Spotted seatrout
# - Red drum
# - White shrimp
# - Brown shrimp
# - Lane snapper

##Ok. We are going to try a couple of things. Here are the different models we are going to try:
#Dependent variable:
# Binary
# Abundance with zeros
# Hurdle RF
#Independent variables:
# All
# Handpicked set
```

## Binary with all variables

This model will translate the abundance data into presence/absence data and then train a RF model for binary on all of the independent variables for each species. Will do sled and trawl separately and then together.

First let's organize the raw data into something we can use

```{r}
#First we have to organize the data. This will include isolating only the species of interest and including all of the zeros
#add in all zeros
#Extract all of the station, site, and sample_period data to combine it together after
stat_dat_sled <- sled %>% dplyr::select(site, station, sample_period, light_attenuation_av:pull_id, station_seagrass_richness:sf_station_sd_below_wt_m) %>% distinct()

stat_dat_trawl <- trawl %>% dplyr::select(tow_id:station_sd_rm_sht_ratio, site, station, sample_period, tow_area:light_attenuation_av, seagrass_richness) %>% distinct()

#add in all zeros
sled_zero <- sled %>% dplyr::select(site, station, sample_period, species_common, species_abundance) %>% 
  dplyr::group_by(site, station, sample_period, species_common) %>% 
  dplyr::summarise(species_abundance = sum(species_abundance, na.rm = TRUE), .groups = "drop") %>%
  # Use complete() to generate all combinations
  complete(
    nesting(site, station, sample_period), # Only crosses combinations that actually exist in your data
    species_common,                        # Crosses the above with all unique species
    fill = list(species_abundance = 0)     # Fills implicitly missing abundances with 0
  )

trawl_zero <- trawl %>% dplyr::select(site, station, sample_period, species_common, species_abundance) %>% 
  dplyr::group_by(site, station, sample_period, species_common) %>% 
  dplyr::summarise(species_abundance = sum(species_abundance, na.rm = TRUE), .groups = "drop") %>%
  # Use complete() to generate all combinations
  complete(
    nesting(site, station, sample_period), # Only crosses combinations that actually exist in your data
    species_common,                        # Crosses the above with all unique species
    fill = list(species_abundance = 0)     # Fills implicitly missing abundances with 0
  )

#recombine the data
sledall <- sled_zero %>% left_join(stat_dat_sled, by = c('site', 'station', 'sample_period'))
trawlall <- trawl_zero %>% left_join(stat_dat_trawl, by = c('site', 'station', 'sample_period'))

#Isolate species of interest and replace NAs with zeros
sledsp <- sledall %>% dplyr::filter(species_common %in% c('gray snapper', 'blue crab', 'pink shrimp', 'white shrimp', 'brown shrimp', 'spotted seatrout', 'lane snapper', 'red drum')) %>% mutate(across(where(is.numeric), ~replace_na(.x, 0))) %>% dplyr::select(-c(spp_x, hd_station_ep_wt_gg, hd_station_sd_ep_wt_gg, hd_station_below_wt_m, hd_station_sd_below_wt_m, spp_y, he_station_sd_ep_wt_gg, he_station_sd_below_wt_m, rm_stationsp_sd_shoots, rm_stationsp_sd_shoots_m, rm_stationsp_se_shoots_m, rm_station_sd_above_wt_m, rm_station_ep_wt_gg, rm_station_sd_ep_wt_gg, rm_station_below_wt_m, rm_station_sd_below_wt_m))

trawlsp <- trawlall %>% dplyr::filter(species_common %in% c('gray snapper', 'blue crab', 'pink shrimp', 'white shrimp', 'brown shrimp', 'spotted seatrout', 'lane snapper', 'red drum')) %>% mutate(across(where(is.numeric), ~replace_na(.x, 0)))

#Ok. We have our dataset with abundances and zeros. We are going to add a column to indicate binary presence/absence and an ID column
sledbin <- sledsp %>% mutate(pa = if_else(species_abundance == 0, 0, 1), ID = row_number())

trawlbin <- trawlsp %>% mutate(pa = if_else(species_abundance == 0, 0, 1), ID = row_number())
```

We set up the binary data for both the sled and trawl data. Let's set up the RF binary model

```{r}
#Going to start with sled data with brown shrimp
data <- sledbin %>% filter(species_common == 'brown shrimp')

data$pa <- as.factor(data$pa)

#Take out columns that will not be in model
data <- data %>% dplyr::select(-c(station, sample_period, species_common, species_abundance, secchi, pull_id, site))

str(data)
data$depth <- as.numeric(data$depth)
data$station_seagrass_richness <- as.numeric(data$station_seagrass_richness)
set.seed(19)
# Randomly select 70% of the data frame for the training dataset
data.train <- data[sample(1:nrow(data), nrow(data) * 0.7, replace = FALSE), ]
# Remainder (i.e., 30%) becomes the test dataset.
data.test <- data[!(data$ID %in% data.train$ID), ]

data.train <- data.train %>% dplyr::select(-ID)
data.test <- data.test %>% dplyr::select(-ID)

#set tasks for training and test datasets
task_train <- as_task_classif(x = data.train, target = 'pa', positive = '1')
task_test <- as_task_classif(x = data.test, target = 'pa', positive = '1')

# Make learner.
learner <-lrn(
  "classif.ranger",
  predict_type = "prob",
  mtry  = to_tune(1, ncol(data.train) - 4),
  sample.fraction = to_tune(0.2, 0.9),
  min.node.size = to_tune(p_int(1,10)),
  importance = 'impurity'
)

#tune hyperparameters
instance = ti(
  task = task_train,
  learner = learner,
  resampling = rsmp("cv", folds = 5),
  measures = msr("classif.ce"),
  terminator = trm("none")
)

tuner = mlr3tuning::tnr("grid_search", resolution = 5, batch_size = 5)
tuner$optimize(instance) # Takes ~ 4 minutes on my relatively fast computer

#store tuned hyperparameters in learner
learner$param_set$values <- instance$result_learner_param_vals

#finally! We can train our model with the train function
learner$train(task_train)

#let's quickly look at the model
learner$model

#Accuracy of model - first on training data, then testing data
measures <- msrs(c('classif.acc'))

pred_train <- learner$predict(task_train)
pred_train$confusion
pred_train$score(measures)


pred_test <- learner$predict(task_test)
pred_test$confusion
pred_test$score(measures)

#importance with iml package - this is looking at the most influencial predictors in the model
x_data <- data.train %>% dplyr::select(-pa) 
# Create "Predictor" object to interpret findings via the iml package.
predictor_data <- Predictor$new(learner, data = x_data, y = data.train$pa) 

imp_data <- FeatureImp$new(predictor_data, loss = "ce") # Calculate importance.

imp_df_data <- imp_data$results 

#so salinity is the most important variable for brown shrimp. Let's investigate how it impacts the presence of shrimp

effect_sal <- FeatureEffect$new(predictor_data, feature = c('salinity'), method = 'pdp')
effect_sal$plot()


```

